### 2020-03-31 SI_code_revised.Rmd
Tables numbering corrected

---
title: Towards more realistic model representation of microbial population dynamics in complex environment
author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Michal Choma, Jianqiu Zheng, Nancy Hess, Kirsten Hofmockel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supplementary Information
## Step-by-step analysis and code

The analysis requires several libraries

```{r libraries loading, message=FALSE, warning=FALSE}

##libraries
library(deSolve)
library(dplyr)
library(FME)
library(reshape)
library(ggplot2)
library(foreach)
library(doParallel)
library(openxlsx)
library(DEoptim)
library(gridExtra)
library(kableExtra)
library(lmerTest)
library(psycho)
library(ggeffects)
library(rcompanion)
library(vegan)

##ggplot theme
theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
                 axis.title=element_text(size=18, colour="black"),
                 axis.line=element_line(size=0.5, colour="black"),
                 strip.text=element_text(size=18, face="bold"),
                 axis.ticks=element_line(size=1, colour="black"),
                 axis.ticks.length=unit(-0.05, "cm"),
                 panel.background=element_rect(colour="black", fill="white"),
                 panel.grid=element_line(linetype=0),
                 legend.text=element_text(size=14, colour="black"),
                 legend.title=element_text(size=14, colour="black"),
                 legend.position=c("right"),
                 legend.key.size=unit(1, "cm"),
                 strip.background=element_rect(fill="grey98", colour="black"),
                 legend.key=element_rect(fill="white", size=1.2),
                 legend.spacing=unit(0.5, "cm"),
                 plot.title=element_text(size=18, face="bold", hjust=-0.05))
```

Data are available at https://github.com/petacapek/MinT.

```{r data loading, echo=FALSE}
#respiration rates in umol/ml/h
resp<-read.csv(file=c("data_checked_respiration_raw.csv"))

#arrange the data
resp.ordered<-resp[order(resp$Sample, resp$Day), ]

#biological data
biology<-read.csv(file=c("data_checked_biology_raw.csv"))

#arrange the data in the same way
biology.ordered<-biology[order(biology$Sample, biology$Day), -c(2,3)]
biology.ordered$Sample<-paste0("CSub3_", biology.ordered$Sample)

#merging both data sets
m0<-merge(resp.ordered, biology.ordered, by.x=c("Sample", "Day"), by.y = c("Sample", "Day"), all=T)

```

The conducted experiment is large. For this analysis, only subset of data is used. We use treatments with Glucose or Celluloze as a single source of organic carbon and all three levels of incubation system complexity (i.e. broth, glass wool and mix of glass beads).

```{r data filtering}

d<-subset(m0, Substrate=="Glucose" | Substrate=="Cellobiose")
summary(d)

#removing outliers
#respiration rate
ggplot(d, aes(Time, r))+geom_point(cex=6)+facet_wrap(Structure~Substrate, scales="free")

d[(d$Substrate=="Glucose" & d$Structure=="Glass wool" & d$Time>20 & d$Time<30 & d$r<0.4 & !is.na(d$r)), "r"]<-NA

#Proteins
ggplot(d, aes(Time, Prot.in))+geom_point(cex=6)+facet_wrap(Structure~Substrate, scales="free")

d[(d$Substrate=="Cellobiose" & d$Structure=="Broth" &  d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Glucose" & d$Structure=="Broth" & d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Cellobiose" & d$Structure=="Mixed glass" & d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Cellobiose" & d$Structure=="Glass wool" & d$Time>75 & d$Prot.in>110 & !is.na(d$Prot.in)), "Prot.in"]<-NA

```

DNA and protein concentrations are in $\mu$grams per incubation vial. Both values are converted to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al. (2004).

```{r conversions}

d$Protc<-d$Prot.in*0.46/12.01/4
d$DNAc<-d$DNA*0.51/12.01/4

```

## Statistical analysis

Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration. Because all three variables display significant variability over time, the time of measurement is treated as a random effect factor. Type of the barrier is nested within the substrate.

```{r lme analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#respiration rate
lme1<-lmer(r~Substrate+Structure/Substrate + (1|Day), d)
#Proteins
lme2<-lmer(Protc~Substrate+Structure/Substrate + (1|Day), d)
#DNA
lme3<-lmer(DNAc~Substrate+Structure/Substrate + (1|Day), d)

t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme2)[c(5,3,6)], anova(lme3)[c(5,3,6)]))

row.names(t.s0)<-c("Substrate", "Structure", "Substrate*Structure")

kable(t.s0, digits = c(2, 1, 3, 2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S1:** Results of linear mixed-effects model analysis. Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for measured respiration rate, cellular protein concentration and DNA. F values, degrees of freedom (df) and the probability (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Respiration rate" = 3, "Cellular Proteins" = 3, "DNA" = 3),
                   italic=T)

```
$~$
$~$
$~$
$~$
$~$
$~$

```{r lme effects, echo=FALSE, fig.height=5, fig.width=15, message=FALSE, paged.print=FALSE, fig.cap="Fig. S1: Marginal effects of the substrate (Glucose or Cellobiose) and the type of physical barrier (BROTH, GLASS, WOOL) on the respiration rate (a), cellular proteins (b) and DNA concentration (c) estimated by the linear mixed-effect model."}
#respiration
lme1_contr<-get_contrasts(lme1)
lme1_effects<-ggpredict(lme1, c("Substrate", "Structure"), type="fe")
lme1_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
lme1_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
lme1_effects<-lme1_effects[order(lme1_effects$group), ]
lme1_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme1_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme1_contr, threshold = 0.05)[[2]])[c(3:4)])
#Proteins
lme2_contr<-get_contrasts(lme2)
lme2_effects<-ggpredict(lme2, c("Substrate", "Structure"), type="fe")
lme2_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
lme2_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
lme2_effects<-lme2_effects[order(lme2_effects$group), ]
lme2_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme2_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme2_contr, threshold = 0.05)[[2]])[c(3:4)])
#DNA
lme3_contr<-get_contrasts(lme3)
lme3_effects<-ggpredict(lme3, c("Substrate", "Structure"), type="fe")
lme3_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
lme3_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
lme3_effects<-lme3_effects[order(lme3_effects$group), ]
lme3_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme3_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme3_contr, threshold = 0.05)[[2]])[c(3:4)])

p1<-ggplot(lme1_effects, aes(x, predicted))+
    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6, show.legend = F)+
    scale_fill_manual(values = c("black", "grey", "white"))+
    scale_colour_manual(values = c("black", "black", "black"))+
    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
                  position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
    ylab(expression(paste("Respiration rate (", mu, "mol",(C-CO[2])~ml^{-1}~h^{-1}, ")")))+
    ggtitle("a)")+
    geom_text(aes(label=letters, colour=group, vjust=-10, fontface="bold.italic"),
              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
    ylim(-0.01,0.3)+
    theme_min+theme(axis.title.x = element_blank(),
                    legend.title = element_blank())
p2<-ggplot(lme2_effects, aes(x, predicted))+
    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6, show.legend = F)+
    scale_fill_manual(values = c("black", "grey", "white"))+
    scale_colour_manual(values = c("black", "black", "black"))+
    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
                  position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
    ylab(expression(paste("Cellular Proteins (", mu, "mol",(C[Proteins])~ml^{-1}, ")")))+
    ggtitle("b)")+
    geom_text(aes(label=letters, colour=group, vjust=-3, fontface="bold.italic"),
              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
    ylim(0,1.2)+
    theme_min+theme(axis.title.x = element_blank(),
                    legend.title = element_blank())
p3<-ggplot(lme3_effects, aes(x, predicted))+
    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6)+
    scale_fill_manual(values = c("black", "grey", "white"))+
    scale_colour_manual(values = c("black", "black", "black"))+
    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
                  position=position_dodge(width = 0.5), width=0.1)+
    ylab(expression(paste("DNA (", mu, "mol",(C[DNA])~ml^{-1}, ")")))+
    ggtitle("c)")+
    geom_text(aes(label=letters, colour=group, vjust=-3, fontface="bold.italic"),
              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
    ylim(0,0.08)+
    theme_min+theme(axis.title.x = element_blank(),
                    legend.title = element_blank(),
                    legend.position = c(0.25,0.2))

grid.arrange(p1, p2, p3, nrow=1)

```
$~$
$~$
$~$

## Microbial-explicit models analysis

### Both models are first calibrated against respiration rate

Calibration of both models is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

The calibration of the models is not run within this document because it takes couple of hours to finish. Results of the previous iteration is printed instead. 
To run the code and reproduce the results, the option "eval" need to be changed to TRUE. Please note that the parallel computing is used. Number of cores is set to total number of cores of the computer minus 1. This can be changed if necessary.

When models are calibrated against the respiration rate only, initial concentration of microbial biomass or reserves and structures is estimated as a single parameter. However, when model parameters are estimated for each treatment separately, initial microbial biomass can differ between treatments. This does not correspond to a reality. Therefore, initial concentration of microbial biomass is subsequently fixed across treatments.

#### Monod model

```{r monod respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monod_r.R")
source("../Rscripts/monod_r_fix.R")
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)
#########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##########################################################
#for different structures
monod_r2<-monod_r(data=d, FACT = 2)
monod_r2_fix<-monod_r_fix(data=d, FACT = 2)
#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)
monod_r3_fix<-monod_r_fix(data=d, FACT = 1)
#for each separately  
monod_r4<-monod_r(data=d, FACT = 3)
monod_r4_fix<-monod_r_fix(data=d, FACT = 3)
###################################################
stopImplicitCluster()
###################################################
# monod_r1$goodness
# monod_r2$goodness
# monod_r3$goodness
# monod_r4$goodness

# mean(c(monod_r4[[1]]$pars[["Cmic_0"]], monod_r4[[2]]$pars[["Cmic_0"]],
#   monod_r4[[3]]$pars[["Cmic_0"]], monod_r4[[4]]$pars[["Cmic_0"]],
#   monod_r4[[5]]$pars[["Cmic_0"]], monod_r4[[6]]$pars[["Cmic_0"]]))

monod_r1$goodness
monod_r2_fix$goodness
monod_r3_fix$goodness
monod_r4_fix$goodness
```

```{r monod respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# monod_r_results_par<-as.data.frame(rbind(monod_r1[[1]]$pars[1:4],
#                       monod_r2_fix[[1]]$pars, monod_r2_fix[[2]]$pars,
#                       monod_r2_fix[[3]]$pars,
#                       monod_r3_fix[[1]]$pars, monod_r3_fix[[2]]$pars,
#                       monod_r4_fix[[1]]$pars, monod_r4_fix[[2]]$pars,
#                       monod_r4_fix[[3]]$pars, monod_r4_fix[[4]]$pars,
#                       monod_r4_fix[[5]]$pars, monod_r4_fix[[6]]$pars))
# monod_r_results$Cmic_init<-c(0.09125118)
# monod_r_results_good<-rbind(monod_r1$goodness,
#                             monod_r2_fix$goodness,
#                             monod_r3_fix$goodness,
#                             monod_r4_fix$goodness)
# write.csv(monod_r_results_par, file = c("../monod_r_results_par.csv"), row.names = F)
# write.csv(monod_r_results_good, file = c("../monod_r_results_good.csv"), row.names = F)

monod_r_results_par<-read.csv("monod_r_results_par.csv")
monod_r_results_good<-read.csv("monod_r_results_good.csv")

t.s1.1<-data.frame(Vmax=monod_r_results_par[,1],
                 Km=monod_r_results_par[,2],
                 CUE=monod_r_results_par[,3],
                 k=monod_r_results_par[,4])
row.names(t.s1.1)<-c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S2:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S3:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

#### Sub-microbial model

```{r Sub-microbial model respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_r.R")
source("../Rscripts/two_pool_r_fix.R")
#across all treatments
two_pool_r1<-two_pool_r(data=d, FACT = 4)
#################################################
no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)
#################################################
#for different structures
two_pool_r2<-two_pool_r(data=d, FACT = 2)
two_pool_r2_fix<-two_pool_r_fix(data=d, FACT = 2)
#for different substrates
two_pool_r3<-two_pool_r(data=d, FACT = 1)
two_pool_r3_fix<-two_pool_r_fix(data=d, FACT = 1)
#for each separately 
two_pool_r4<-two_pool_r(data=d, FACT = 3)
two_pool_r4_fix<-two_pool_r_fix(data=d, FACT = 3)
#################################################
stopImplicitCluster()
#################################################
# mean(c(two_pool_r4[[1]]$pars[["R_0"]], two_pool_r4[[2]]$pars[["R_0"]],
#      two_pool_r4[[3]]$pars[["R_0"]], two_pool_r4[[4]]$pars[["R_0"]],
#      two_pool_r4[[5]]$pars[["R_0"]], two_pool_r4[[6]]$pars[["R_0"]]))
# mean(c(two_pool_r4[[1]]$pars[["S_0"]], two_pool_r4[[2]]$pars[["S_0"]],
#      two_pool_r4[[3]]$pars[["S_0"]], two_pool_r4[[4]]$pars[["S_0"]],
#      two_pool_r4[[5]]$pars[["S_0"]], two_pool_r4[[6]]$pars[["S_0"]]))

two_pool_r1$goodness
two_pool_r2_fix$goodness
two_pool_r3_fix$goodness
two_pool_r4_fix$goodness

```

```{r Sub-microbial model respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:5)],
#                    two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
#                    two_pool_r2_fix[[3]]$pars,
#                    two_pool_r3_fix[[1]]$pars, two_pool_r3_fix[[2]]$pars,
#                    two_pool_r4_fix[[1]]$pars, two_pool_r4_fix[[2]]$pars,
#                    two_pool_r4_fix[[3]]$pars, two_pool_r4_fix[[4]]$pars,
#                    two_pool_r4_fix[[5]]$pars, two_pool_r4_fix[[6]]$pars)
# two_pool_r_results_good<-rbind(two_pool_r1$goodness,
#                               two_pool_r2$goodness,
#                               two_pool_r3$goodness,
#                               two_pool_r4$goodness)
# write.csv(two_pool_r_results_par, file = c("./two_pool_r_results_par.csv"), row.names = F)
# write.csv(two_pool_r_results_good, file = c("./two_pool_r_results_good.csv"), row.names = F)

two_pool_r_results_par<-read.csv("two_pool_r_results_par.csv")
two_pool_r_results_good<-read.csv("two_pool_r_results_good.csv")

t.s3.1<-data.frame(Vmax=two_pool_r_results_par[,1],
                 Km=two_pool_r_results_par[,2],
                 f=two_pool_r_results_par[,3],
                 k=two_pool_r_results_par[,4],
                 Yu=two_pool_r_results_par[,5])
row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S4:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S5:** Goodness of fit of Sub-microbial model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

### Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.

Both models are calibrated against all data we measured. In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated. For Monod model, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass. For Sub-microbial model, initial concentration of Structures is defined similarly - i.e. initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures. However, initial concetration of Reserves cannot be defined because initial concentration of proteins is not know. Therefore, one more parameter is defined. This parameter is denoted as RS_0 in the respective function and define initial concentration of Reserves per unit of Structures. This parameter has to be fixed across treatments to reflect the reality.
In addition, Monod model is calibrated against respiration rate and proteins or DNA concentration separately. When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.

#### Monod model

```{r Monod model - biomass components, eval=F, echo=T}
#Proteins
#Loading the function 
source("../Rscripts/monod_prot.R")
source("../Rscripts/monod_prot_fix.R")
#across all treatments
monod_prot1<-monod_prot(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_prot2<-monod_prot(data=d, FACT = 2)
monod_prot2_fix<-monod_prot_fix(data=d, FACT = 2)
#for different substrates
monod_prot3<-monod_prot(data=d, FACT = 1)
monod_prot3_fix<-monod_prot_fix(data=d, FACT = 1)
#for each separately 
monod_prot4<-monod_prot(data=d, FACT = 3)
monod_prot4_fix<-monod_prot_fix(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_prot1$goodness
monod_prot2_fix$goodness
monod_prot3_fix$goodness
monod_prot4_fix$goodness

###############################################################
#DNA
#Loading the function 
source("../Rscripts/monod_i_DNA.R")
#across all treatments
monod_DNA1<-monod_DNA(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_DNA2<-monod_DNA(data=d, FACT = 2)
#for different substrates
monod_DNA3<-monod_DNA(data=d, FACT = 1)
#for each separately 
monod_DNA4<-monod_DNA(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_DNA1$goodness
monod_DNA2$goodness
monod_DNA3$goodness
monod_DNA4$goodness
###############################################################
#Proteins and DNA
#Loading the function 
source("../Rscripts/monod_all.R")
#across all treatments
monod_all1<-monod_all(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_all2<-monod_all(data=d, FACT = 2)
#for different substrates
monod_all3<-monod_all(data=d, FACT = 1)
#for each separately 
monod_all4<-monod_all(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_all1$goodness
monod_all2$goodness
monod_all3$goodness
monod_all4$goodness
```

```{r Monod overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# #results of the previous iteration (not run)
# ##All measurements together
# monod_all_results_par<-rbind(monod_all1[[1]]$pars, 
#                              monod_all2[[1]]$pars, monod_all2[[2]]$pars,
#                              monod_all2[[3]]$pars, 
#                              monod_all3[[1]]$pars, monod_all3[[2]]$pars,
#                              monod_all4[[1]]$pars, monod_all4[[2]]$pars,
#                              monod_all4[[3]]$pars, monod_all4[[4]]$pars,
#                              monod_all4[[5]]$pars, monod_all4[[6]]$pars)
# 
# monod_all_results_good<-rbind(monod_all1$goodness,
#                               monod_all2$goodness,
#                               monod_all3$goodness,
#                               monod_all4$goodness)
# 
# write.csv(monod_all_results_par, file = c("../monod_all_results_par.csv"), row.names = F)
# write.csv(monod_all_results_good, file = c("../monod_all_results_good.csv"), row.names = F)
# 
# ##Respiration rate and proteins
# monod_prot_results_par<-rbind(monod_prot1[[1]]$pars[c(1:5)], 
#                              monod_prot2_fix[[1]]$pars, monod_prot2_fix[[2]]$pars,
#                              monod_prot2_fix[[3]]$pars, 
#                              monod_prot3_fix[[1]]$pars, monod_prot3_fix[[2]]$pars,
#                              monod_prot4_fix[[1]]$pars, monod_prot4_fix[[2]]$pars,
#                              monod_prot4_fix[[3]]$pars, monod_prot4_fix[[4]]$pars,
#                              monod_prot4_fix[[5]]$pars, monod_prot4_fix[[6]]$pars)
# 
# monod_prot_results_good<-rbind(monod_prot1$goodness,
#                              monod_prot2_fix$goodness,
#                              monod_prot3_fix$goodness,
#                              monod_prot4_fix$goodness)
# 
# write.csv(monod_prot_results_par, file = c("../monod_prot_results_par.csv"), row.names = F)
# write.csv(monod_prot_results_good, file = c("../monod_prot_results_good.csv"), row.names = F)
# 
# ##Respiration rate and DNA
# monod_DNA_results_par<-rbind(monod_DNA1[[1]]$pars, 
#                              monod_DNA2[[1]]$pars, monod_DNA2[[2]]$pars,
#                              monod_DNA2[[3]]$pars, 
#                              monod_DNA3[[1]]$pars, monod_DNA3[[2]]$pars,
#                              monod_DNA4[[1]]$pars, monod_DNA4[[2]]$pars,
#                              monod_DNA4[[3]]$pars, monod_DNA4[[4]]$pars,
#                              monod_DNA4[[5]]$pars, monod_DNA4[[6]]$pars)
# 
# monod_DNA_results_good<-rbind(monod_DNA1$goodness,
#                               monod_DNA2$goodness,
#                               monod_DNA3$goodness,
#                               monod_DNA4$goodness)
# 
# write.csv(monod_DNA_results_par, file = c("../monod_DNA_results_par.csv"), row.names = F)
# write.csv(monod_DNA_results_good, file = c("../monod_DNA_results_good.csv"), row.names = F)

#All measurements
monod_all_results_par<-read.csv("monod_all_results_par.csv")
monod_all_results_good<-read.csv("monod_all_results_good.csv")
monod_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
monod_all_results_good[monod_all_results_good$R2<0, "R2"]<-0
#Respiration and proteins
monod_prot_results_par<-read.csv("monod_prot_results_par.csv")
monod_prot_results_good<-read.csv("monod_prot_results_good.csv")
monod_prot_results_good$variable<-rep(c("Respiration", "Proteins"), 4)
monod_prot_results_good[monod_prot_results_good$R2<0, "R2"]<-0
#Respiration and DNA
monod_DNA_results_par<-read.csv("monod_DNA_results_par.csv")
monod_DNA_results_good<-read.csv("monod_DNA_results_good.csv")
monod_DNA_results_good$variable<-rep(c("Respiration", "DNA"), 4)
monod_DNA_results_good[monod_DNA_results_good$R2<0, "R2"]<-0

t.s4.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monod_all_results_par[,1],
                   Km=monod_all_results_par[,2],
                   CUE=monod_all_results_par[,3],
                   k=monod_all_results_par[,4],
                   fp=monod_all_results_par[,5],
                   fd=monod_all_results_par[,6])

options(knitr.kable.NA = '')

kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S6:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S7:** Goodness of fit of Monod model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

t.s4.2<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monod_prot_results_par[,1],
                   Km=monod_prot_results_par[,2],
                   CUE=monod_prot_results_par[,3],
                   k=monod_prot_results_par[,4],
                   fp=monod_prot_results_par[,5])

options(knitr.kable.NA = '')

kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S8:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S9:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

t.s4.3<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monod_DNA_results_par[,1],
                   Km=monod_DNA_results_par[,2],
                   CUE=monod_DNA_results_par[,3],
                   k=monod_DNA_results_par[,4],
                   fd=monod_DNA_results_par[,5])

options(knitr.kable.NA = '')

kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S10:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S11:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

```

#### Sub-microbial model

Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.

```{r Sub-microbial model calibration, eval=F, echo=T}
#Loading the function 
source("./Rscripts/two_pool_allx.R")
source("../Rscripts/two_pool_allx_fix.R")
#across all treatments
two_pool_allx_out1<-two_pool_allx(data=d, FACT = 4)

###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
two_pool_allx_out2<-two_pool_allx(data=d, FACT = 2)

two_pool_allx_fix_out2<-two_pool_allx_fix(data=d, FACT = 2)
two_pool_allx_fix_out2$goodness

two_pool_allx_fix_out2[[1]]$pars
two_pool_allx_fix_out2[[2]]$pars
two_pool_allx_fix_out2[[3]]$pars

# mean(c(two_pool_allx_fix_out2[[1]]$pars[["Vmax"]],
#        two_pool_allx_fix_out2[[2]]$pars[["Vmax"]],
#        two_pool_allx_fix_out2[[3]]$pars[["Vmax"]]))
# mean(c(two_pool_allx_fix_out2[[1]]$pars[["Yu"]],
#        two_pool_allx_fix_out2[[2]]$pars[["Yu"]],
#        two_pool_allx_fix_out2[[3]]$pars[["Yu"]]))
# mean(c(two_pool_allx_fix_out2[[1]]$pars[["k"]],
#        two_pool_allx_fix_out2[[2]]$pars[["k"]],
#        two_pool_allx_fix_out2[[3]]$pars[["k"]]))
############################################################################
#in the following code, the Vmax, Ys and k parameters are fixed across treatments
#becuase its variability across treatments is minimal (data not shown)
source("./Rscripts/two_pool_all_fixVmaxYuk.R")
two_pool_all_fixVmaxYuk_out2<-two_pool_all_fixVmaxYuk(data=d, FACT = 2)
two_pool_all_fixVmaxYuk_out2$goodness

############################################################################
#for different substrates
two_pool_allx_fix_out3<-two_pool_allx_fix(data=d, FACT = 1)
two_pool_allx_fix_out3$goodness
#for each separately 
two_pool_allx_fix_out4<-two_pool_allx_fix(data=d, FACT = 3)
two_pool_allx_fix_out4$goodness
#in the following code, the Vmax Ys, and k parameters are fixed across treatments
two_pool_all_fixVmaxYuk_out4<-two_pool_all_fixVmaxYuk(data=d, FACT = 3)
two_pool_all_fixVmaxYuk_out4$goodness
###########################################################
stopImplicitCluster()
###########################################################
two_pool_allx_out1$goodness
two_pool_all_fixVmaxYuk_out2$goodness
two_pool_allx_fix_out3$goodness
two_pool_all_fixVmaxYuk_out4$goodness

# mean(c(two_pool_allx_out4[[1]]$pars[["RS_0"]],
#        two_pool_allx_out4[[2]]$pars[["RS_0"]],
#        two_pool_allx_out4[[3]]$pars[["RS_0"]],
#        #two_pool_allx_out4[[4]]$pars[["RS_0"]],
#        two_pool_allx_out4[[5]]$pars[["RS_0"]],
#        two_pool_allx_out4[[6]]$pars[["RS_0"]]))

```

```{r Sub-microbial model overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# two_pool_all_results_par<-rbind(c(two_pool_allx_out1[[1]]$pars[c(1:8)], RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out2[[1]]$pars[1],
#                           two_pool_all_fixVmaxYuk_out2[[1]]$pars[2],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out2[[1]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out2[[2]]$pars[1],
#                           two_pool_all_fixVmaxYuk_out2[[2]]$pars[2],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out2[[2]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out2[[3]]$pars[1],
#                           two_pool_all_fixVmaxYuk_out2[[3]]$pars[2],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out2[[3]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(two_pool_allx_fix_out3[[1]]$pars[c(1:8)], RS_0 = 0.04012019),
#                         c(two_pool_allx_fix_out3[[2]]$pars[c(1:8)], RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out4[[1]]$pars[c(1:2)],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out4[[1]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out4[[2]]$pars[c(1:2)],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out4[[2]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out4[[3]]$pars[c(1:2)],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out4[[3]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out4[[4]]$pars[c(1:2)],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out4[[4]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out4[[5]]$pars[c(1:2)],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out4[[5]]$pars[c(3:5)],
#                           RS_0 = 0.04012019),
#                         c(Vmax = 0.2468173, 
#                           two_pool_all_fixVmaxYuk_out4[[6]]$pars[c(1:2)],
#                           k = 0.007046304,
#                           Yu = 0.3731948,
#                           two_pool_all_fixVmaxYuk_out4[[6]]$pars[c(3:5)],
#                           RS_0 = 0.04012019)
#                         )
# two_pool_all_results_good<-rbind(two_pool_allx_out1$goodness,
#                                two_pool_all_fixVmaxYuk_out2$goodness,
#                                two_pool_allx_fix_out3$goodness,
#                                two_pool_all_fixVmaxYuk_out4$goodness)
# 
# write.csv(two_pool_all_results_par, file = c("./two_pool_all_results_par.csv"), row.names = F)
# write.csv(two_pool_all_results_good, file = c("./two_pool_all_results_good.csv"), row.names = F)

two_pool_all_results_par<-read.csv("two_pool_all_results_par.csv")
two_pool_all_results_good<-read.csv("two_pool_all_results_good.csv")
two_pool_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
two_pool_all_results_good[two_pool_all_results_good$R2<0, "R2"]<-0

t.s6.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=two_pool_all_results_par[,1],
                   Km=two_pool_all_results_par[,2],
                   f=two_pool_all_results_par[,3],
                   k=two_pool_all_results_par[,4],
                   Yu=two_pool_all_results_par[,5],
                   fpr=two_pool_all_results_par[,6],
                   fps=two_pool_all_results_par[,7],
                   fds=two_pool_all_results_par[,8],
                   RS_0=two_pool_all_results_par[,9])

options(knitr.kable.NA = '')

kable(t.s6.1, digits = c(0, 2, 2, 2, 3, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S12:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S13:** Goodness of fit of Sub-microbial model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

```